\protect Connecting every bit of knowledge: \\ The structure of Wikipedia's First Link Network

### 2016-09-07 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
typo
citation for xefer added

By following the first hyperlink contained in all articles of the English edition of
Wikipedia as of November, 2014, we form a directed network:
{\it Wikipedia's First Link Network} (FLN).
Inspired by the claim that the majority of first links lead to
``Philosophy''---popularized by an 
\href{https://xkcd.com/903/}{xkcd}
comic and subsequently
discussed in blog posts 
\cite{xkcd, mat_blog, Ilmari_first_links, 
bob_west}
---
we holistically study how the FLN  yields a flow of connections.
For methodological details, see Sec.~\ref{Appendix}.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{centrality_measures.pdf}
  \caption{
    \textbf{Centrality Measures}
    The table above compares standard centrality measures on the sample network against traversal funnels.
    }
  \label{fig:Centrality Measures}

@@

To compute traversal funnels on a directed network, begin at each node: 1. Traverse the edge to the next node 2. Increment the next node's count 3. Repeat while the next node is unvisited. The resulting count is each node's number of traversal funnels: the number of paths the node directs up to the start of a cycle.

@@

For our exploration of the FLN, by considering the FLN not only as collection of directly linked pairs of articles, but
as a flow, with accumulated references, cycles, and path-connected groups, we can explore how the many articles in Wikipedia are organized and related.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{popularity_visits.pdf}
  \caption{
\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal visits.}
We use the total number of page views provided by Wikipedia for the month
of October 2015 to compare each article's popularity to traversal visits.
While the most popular articles do not necessarily correspond to the articles
with the highest number of traversal visits, the variation in popularity appears to decrease as the number of traversal visits increases. The greatest number of articles fall within roughly 1 million traversal visits and page views. Overall, the page views for the top 1000 articles articles by traversal visits appear to be log-normally distributed. 
}
  \label{fig:Views and Visits}

\end{figure}
As a distribution, we find few articles influence the structure of the 
FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
more than $99\%$ with none---most articles are recipients of 
the references flowing through the articles with at least one traversal funnel.
When to a power law 
    distribution (using the Maximum Likelihood Estimator) with xmin as the number of traversal funnels minimizing Kolmogorov-Smirnov distance), we find xmin to be 30 with a corresponding power law rank exponent $\alpha \simeq 1.73$ and a size exponent $\gamma \simeq 1.58$
(see Fig.~\ref{fig:Funnels Distribution}). 
When compared against alternative distributions, we find inconclusive results for lognormal and truncated power law distributions and significant log likelihood ratios favoring a power law for exponential ($R>0, p=5.3\times10^{-6}$) and stretched exponential ($R>0, p = 1.8\times10^{-8}$) distributions.
Even within the few articles
influencing the structure of the FLN, only a handful of these exert most of the 
control.

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{popularity_funnels.pdf}
  \caption{
\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal funnels.}
We use the total number of page views provided by Wikipedia for the month
of October 2015 to compare each article's popularity to traversal funnels.
}
  \label{fig:Views and Funnels}
\end{figure}
\subsection{Article Popularity}

@@

>>>>

By following the first hyperlink contained in all articles of the English edition of
Wikipedia as of November, 2014, we form a directed network:
{\it Wikipedia's First Link Network} (FLN).
Inspired by the claim that the majority of first links lead to
``Philosophy''---popularized by an 
\href{https://xkcd.com/903/}{xkcd}
comic and subsequently
discussed in blog posts 
\cite{xkcd, mat_blog, Ilmari_first_links, 
bob_west, xefer2011philosophy}
---
we holistically study how the FLN  yields a flow of connections.
For methodological details, see Sec.~\ref{Appendix}.

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{centrality_measures.pdf}
  \caption{
    \textbf{Centrality Measures}
    The table above compares standard centrality measures on the sample network against traversal funnels. While betweenness centrality and eigenvector centrality both reward node A for its place in the cycle, traversal funnels only rewards node C for directing paths into the cycle.
    }
  \label{fig:Centrality Measures}

@@

To compute traversal funnels on a directed network, begin at each node: 1. Traverse the edge to the next node 2. Increment the next node's count by one 3. Repeat while the next node is unvisited. The resulting count is each node's number of traversal funnels: the number of paths the node directs up to the start of a cycle.

@@

By considering the FLN not only as collection of directly linked pairs of articles, but
as a flow, with accumulated references, cycles, and path-connected groups, we explore how the many articles in Wikipedia are organized and related.

@@

As a distribution, we find few articles influence the structure of the 
FLN. Only $17, 821$ articles have one or more traversal funnels, leaving
more than $99\%$ with none---most articles are recipients of 
the references flowing through the articles with at least one traversal funnel.
When to a power law 
    distribution (using the Maximum Likelihood Estimator) with xmin as the number of traversal funnels minimizing Kolmogorov-Smirnov distance), we find xmin to be 30 with a corresponding power law rank exponent $\alpha \simeq 1.73$ and a size exponent $\gamma \simeq 1.58$
(see Fig.~\ref{fig:Funnels Distribution}). 
When compared against alternative distributions, we find inconclusive results for lognormal and truncated power law distributions and significant log likelihood ratios favoring a power law for exponential ($R>0, p=5.3\times10^{-6}$) and stretched exponential ($R>0, p = 1.8\times10^{-8}$) distributions.
Even within the few articles
influencing the structure of the FLN, only a handful of these exert most of the 
control.

\subsection{Article Popularity}

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{popularity_visits.pdf}
  \caption{
\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal visits.}
We use the total number of page views provided by Wikipedia for the month
of October 2015 to compare each article's popularity against traversal visits.
Overall, more page views do not correspond to higher traversal visits---Wikipedia users do not tend to visit articles with a higher accumulation in the FLN more so than they visit others. 
However, the variation in popularity appears to decrease as the number of traversal visits increases. The greatest number of articles fall within roughly 1 million traversal visits and page views. Furthermore, page views for the top 1000 articles articles by traversal visits appear to be log-normally distributed. 
}
  \label{fig:Views and Visits}

\end{figure}
\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{popularity_funnels.pdf}
  \caption{
\textbf{Article Popularity (by page views) for the highest ranking 1000 articles by traversal funnels.}
We use the total number of page views provided by Wikipedia for the month
of October 2015 to compare each article's popularity to traversal funnels.
Similar to traversal visits, more page views do not correspond to higher traversal funnels---Wikipedia users do not tend to visit articles with a higher influence in the FLN more so than they visit others.
}
  \label{fig:Views and Funnels}
\end{figure}

@@

Overall, the contrast between page views and traversal visits or funnels highlights a difference between popularity and influence in the FLN. Although traversal funnels and visits point to influential articles where many topics culminate, Wikipedia users do not visit these articles more than others. What is popular, is not necessarily foundational, and vice versa.
